\bf Supervised tensor decomposition with features on multiple modes

### 2021-06-21 tensor_regress/JCGS/Revision/manuscript/response.tex

    \item \textit{The maximum norm constraint on the linear predictor Theta is a bit ad hoc. How to choose the upper bound $\alpha$ in practice? How does the value of alpha affect the final results? }

    \textbf{Response:} The maximum norm constraint is a technical condition purely for theoretical analysis. In practice, no choice of $\alpha$ is needed.

    We have added the following explanations in Section~\ref{sec:rankM}:
    \begin{quote}
      ``...The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to ensures the existence (boundedness) of MLE. The condition precludes the ill-defined MLE when the optimizer of~\eqref{eq:MLE} diverges to $\pm \infty$; this phenomenon may happen in logistic regression when the Bernoulli responses $\{0,1\}$ are perfectly separable by covariates~\citep{wang2020learning}. For Gaussian models, no maximum norm constraint is needed. In Section~\ref{subsec:statprob}, we show that setting $\alpha$ to an extremely large constant does not compromise the statistical rate in quantities of interest. In practice, the unbounded search is often indistinguishable from the bounded search, since the boundary constraint $\onorm{\Theta}_{\infty} \leq \alpha$ would likely never be active. Similar techniques are commonly used in high-dimensional non-Gaussian problems~\citep{wang2020learning,han2020optimal}....
    \end{quote}

    \item \textit{The authors propose to use BIC to select ranks via grid search. However, for a high-order tensor, one may have to search over a multi-dimensional grid to find the best rank combination. It seems computationally prohibitive (even for a 3-way tensor). Is there any practical guidance about how to more efficiently select the ranks? What are the computational costs for rank selection in the numerical examples? }

    \textbf{Response:} We utilize parallel computing in BIC search to reduce the computational cost. In the absence of parallel computation, one may speed up the rank search using uni-modal properties of marginal BIC. Specifically, we suggest alternating the search between modes and bracket the possible rank range in one mode while holding others fixed. Once the search region is reduced, a small-scale grid search is performed to identify the minimal BIC. We find such practice gives reasonable rank estimates in the two applications considered.

    We have added the following sentence in  Section~\ref{sec:simulation}:
    \begin{quote}
     ``(In simulation)...For each combination, we minimize BIC using a grid search from $(r_1-3, r_2-3, r_3-3)$ to $(r_1+3, r_2+3, r_3+3)$. We removed invalid rank combinations such as $r^2_{\max} \geq \prod_k r_k$ and used parallel search to reduce the computational cost...''
    \end{quote}

    The BIC selection on real data is detailed in Supplementary Notes, Section~\ref{sec:adata}:
    \begin{quote}
         ``(In real data)...The running time for the rank selection via grid search is 95 secs in total, on an iMac macOS High Sierra 10.13.6 with Intel Core i5 3.8 GHz CPU and 8 GB RAM. This indicates the BIC is feasible in the considered setting. "
    \end{quote}

    \item \textit{ In numerical studies, the authors compare the proposed method with a multiple-response GLM approach and comment that the latter is suboptimal because it cannot account for the correlation among responses. There are methods in the literature that can address the issue \citep{yee2003reduced, luo2018leveraging}. It would be interesting to see some updated comparisons. }

    \textbf{Response:} We have added new numerical comparisons using aforementioned references in the current version. See our response to {\bf \# 1, Major Comments Summarized by Editor}.

    \item \textit{In the HCP example, how are the networks constructed in Fig 6? Will the connectivity change with rotations in the equivalence class? In the Nations data example, what would the clustering results be without accounting for the national features (i.e., directly applying Tucker decomposition to the data)? More clarifications are warranted. }

    \textbf{Response:} Figure~\ref{fig:brain} plots top entries in $\tB_{..i}$ for $i=1,2,3,4$. The connectivity is invariant to the rotation of factor matrices. See our response to {\bf \# 6, Reviewer 1}.

    As suggested by reviewer, we have added comparison between supervised vs.\ unsupervised decomposition for  Nations data analysis. The full results are in Section~\ref{sec:compare1}-\ref{sec:compare2} Supplementary Notes:
    \begin{quote}
 ``...We compare the supervised vs.\ unsupervised decomposition in the {\it Nations} data analysis. Table~\ref{tab:unsup} shows the clustering results based on classical unsupervised Tucker decomposition without the feature matrices. Table~\ref{tab:s1} shows the clustering results based on supervised tensor decomposition ({\bf STD}). Compared with supervised decomposition, the unsupervised clustering loses some interpretation. Similar relations {\it exports} and {\it relexports}, {\it ngo} and {\it relngo} are separated into different clusters...''

\begin{table}[!h]
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{c|l}
     Cluster& \multicolumn{1}{c}{Relations}\\
    \hline
    \multirow{2}{*}{ I}  &\textcolor{YellowOrange}{economicaid}, \textcolor{YellowOrange}{releconomicaid}, \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{OliveGreen}{weightedunvote}, \textcolor{OliveGreen}{unweightedunvote}, \\
    &\textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{militaryalliance}\\
    \hline
    \multirow{2}{*}{ II}  & \textcolor{OliveGreen}{warning}, \textcolor{OliveGreen}{violentactions}, \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{duration}, \textcolor{OliveGreen}{severdiplomatic}, \textcolor{OliveGreen}{expeldiplomats}, \textcolor{YellowOrange}{boycottembargo}, \textcolor{BrickRed}{aidenemy}, \\
    &\textcolor{OliveGreen}{negativecomm}, \textcolor{Blue}{accusation}, \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{unoffialacts}, \textcolor{BrickRed}{attackembassy}, \textcolor{Blue}{relemigrants}, \textcolor{OliveGreen}{timesincewar}, \textcolor{BrickRed}{lostterritory}, \textcolor{OliveGreen}{dependent}\\
    \hline
    III & \textcolor{OliveGreen}{timesinceally}, \textcolor{OliveGreen}{independence}, \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{blockpositionindex}\\
    \hline
    \multirow{3}{*}{IV}  &\textcolor{YellowOrange}{treaties}, \textcolor{YellowOrange}{reltreaties}, \textcolor{YellowOrange}{officialvisits}, \textcolor{YellowOrange}{conferences}, \textcolor{YellowOrange}{booktranslations}, \textcolor{YellowOrange}{relbooktranslations}\\
    &\textcolor{OliveGreen}{negativebehavior}, \textcolor{Blue}{nonviolentbehavior}, \textcolor{Blue}{emigrants}, \textcolor{Blue}{emigrants3}, \textcolor{Blue}{students}, \textcolor{Blue}{relstudents}, \textcolor{YellowOrange}{relexports}, \textcolor{YellowOrange}{exports3}\\
    & \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo}, \textcolor{Blue}{intergovorgs3}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{commonbloc1}, \textcolor{OliveGreen}{commonbloc2}\\
    \hline
    \multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
    \end{tabular}
    }
    \caption{Clustering of relations based on unsupervised tensor decomposition. }\label{tab:unsup}
    \end{table}

\begin{table}[!h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{c|l}
 Category & \multicolumn{1}{c}{Relations}\\
\hline
\multirow{2}{*}{ I}   & \textcolor{OliveGreen}{warning}, \textcolor{OliveGreen}{violentactions}, \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{duration}, \textcolor{OliveGreen}{negativebehavior}, \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{severdiplomatic} \\
&\textcolor{OliveGreen}{timesincewar}, \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{commonbloc1}, \textcolor{OliveGreen}{blockpositionindex},  \textcolor{OliveGreen}{expeldiplomats}\\   
\hline
\multirow{2}{*}{II}& \textcolor{Blue}{emigrants}, \textcolor{Blue}{emigrants3}, \textcolor{Blue}{relemigrants}, \textcolor{Blue}{accusation}, \textcolor{Blue}{nonviolentbehavior}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{commonbloc2}, \textcolor{Blue}{intergovorgs3}\\      
&\textcolor{YellowOrange}{releconomicaid}, \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo},  \textcolor{Blue}{students}, \textcolor{Blue}{relstudents}, \textcolor{YellowOrange}{economicaid}, \textcolor{OliveGreen}{negativecomm}, \textcolor{OliveGreen}{militaryalliance} \\
\hline
\multirow{3}{*}{ III} & \textcolor{YellowOrange}{treaties}, \textcolor{YellowOrange}{reltreaties}, \textcolor{YellowOrange}{officialvisits}, \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{YellowOrange}{booktranslations}, \textcolor{YellowOrange}{relbooktranslations}\\
&\textcolor{YellowOrange}{boycottembargo}, \textcolor{OliveGreen}{weightedunvote}, \textcolor{OliveGreen}{unweightedunvote}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{YellowOrange}{exports3}  \\
&\textcolor{YellowOrange}{relexports}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{timesinceally}, \textcolor{OliveGreen}{independence}, \textcolor{YellowOrange}{conferences}, \textcolor{OliveGreen}{dependent}\\
\hline
IV &\textcolor{BrickRed}{aidenemy}, \textcolor{BrickRed}{lostterritory}, \textcolor{OliveGreen}{unoffialacts}, \textcolor{BrickRed}{attackembassy}\\
\hline
\multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
\end{tabular}
}
\caption{Clustering of relations based on supervised tensor decomposition. }\label{tab:s1}
\end{table}

\end{quote}

We comment that, although both methods return clustering outputs, the unsupervised and supervised decomposition tackle different problems. The unsupervised decomposition identifies factors that explain most variation in the tensor, whereas the supervised decomposition identifies factors that are most attributable to side features. There is in general no particular pattern one could expect between these two methods. We provide a simple example here for illustration.
\begin{example}[Complementary information in supervised vs.\ unsupervised factors]
Consider the following data tensor $\tY$ and one-sided feature matrix $\mX$, 
\begin{align}
    \tY = \me_1\otimes \me_1 \otimes \me_1+10\me_2\otimes \me_2 \otimes \me_2, \quad \mX=\me_1,
\end{align}
where $\me_i=(0,\ldots,0, 1,0,\ldots,0)^T$ is the $i$th canonical basis vector in $\mathbb{R}^d$ for $i=1,2$. 
Now, consider the unsupervised vs.\ supervised decomposition of $\tY$ with rank $\mr = (1,1,1)$. 
Then, the top supervised and unsupervised factors are perpendicular to each other, 
\[
\mM_{\text{sup},k} \perp \mM_{\text{unsup},k}, \quad \text{for all } k=1,2,3,
\]
where $\mM_{\text{sup},k}$, $\mM_{\text{unsup},k}$ denote the mode-$k$ factors from supervised and unsupervised decompositions, respectively. 
\end{example}

\begin{rmk}
This example shows complementary information between factors from supervised vs.\ unsupervised decompositions. In general, one could construct examples such that these two methods return {\bf arbitrarily different} factors.   
\end{rmk}

\end{enumerate}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}
